---
title: "Air Quality in New York"
author: "Akash Mer"
date: "`r Sys.Date()`"
output:
    ioslides_presentation:
        keep_md: true
        smaller: true
knit: (function(input, ...){
    rmarkdown::render(input,
        output_dir = "C:/Users/akash/Documents/datasciencecoursera/NY_Air_Quality/docs",
        output_file = file.path("./index"))
    })
---

## **Introduction**

**Exploratory Data Analysis** of daily air quality in New York, from May to September, 1973  
**Code book** : As per the help page of *airquality* data set from the R *datasets* package
  
- **Ozone** - Mean ozone in *parts per billion(ppb)* from 1 to 3 pm at Roosevelt Island
- **Solar.R** - Solar radiation in *Langleys(lang) in the frequency band 4000-7000 Angstroms* from 8 am to 12 pm at Central Park
- **Wind** - Average wind speed in *miles per hour(mph)* from 7 to 10 am at LaGuardia Airport
- **Temp** - Maximum daily temperature in *degrees Fahrenheit($^\circ$F)* at LaGuardia Airport
- **Month** - Month for the measurement date
- **Day** - Day of the month for the measurement date
  
```{r loadData, echo=TRUE, cache=TRUE, message=FALSE, warning=FALSE}
if(system.file(package = "datasets") == "") install.packages("datasets")
library(datasets)
data("airquality")
```

## **Data Processing**
  
`r round((1-mean(complete.cases(airquality)))*100, 2)`% of rows have missing values and these values do have a discernable pattern, as most of them are missing in the month of June, thus they are of the type **MAR**  
These will be imputed with the *impute.knn* function from the **impute** package with the number of neighbours to be used for *imputation = 5 and rng seed set to 325*

```{r transformData, echo=TRUE, cache=TRUE, dependson="loadData"}
if(system.file(package = "impute") == "") install.packages("impute")
library(impute)
dat <- as.data.frame(impute.knn(as.matrix(airquality),k=5,rng.seed = 325)$data)
# Converting the ozone imputed to whole numbers since it is represented in ppm
dat$Ozone <- round(dat$Ozone)
```
  
Thus now the there are no missing values in the data set with the **mean of complete cases in the data set being `r mean(complete.cases(dat))`**

## **Trend of Ozone vs Solar Radiation in NY**

```{r plot1, echo=FALSE, cache=TRUE, fig.width=8, fig.height=5, fig.align='left', warning=FALSE, message=FALSE, dependson=c("loadData", "transformData")}
if(system.file(package = "ggplot2") == "") install.packages("ggplot2")
library(ggplot2)
if(system.file(package = "plotly") == "") install.packages("plotly")
library(plotly)

# Cleaning up the Month column to display the name of the month
dat$Month <- factor(dat$Month, labels = c("May", "June", "July", "August",
                                          "September"))

# Creating the plot with the ggplot
plot1 <- ggplot(dat, aes(Solar.R, Ozone, color = Month)) +
    geom_point() +
    facet_wrap("Month", nrow = 1) +
    scale_color_brewer(palette = "Set2") +
    labs(x = "Solar Radiation(lang)", y = "Ozone (ppb)",
         color = "",
         caption = "Reference: airquality data set from the R datasets package") +
    theme_bw(base_size = 15) +
    theme(legend.position = "none",
          plot.caption = element_text(face = "italic"))

# Displaying the plot as a plotly interactive object
ggplotly(plot1)
```

## **Trend of Ozone vs Wind in NY**

```{r plot2, echo=FALSE, cache=TRUE, fig.width=8, fig.height=5, fig.align='left', warning=FALSE, message=FALSE, dependson=c("loadData", "transformData")}
# Creating the plot with the ggplot
plot2 <- ggplot(dat, aes(Wind, Ozone, color = Month)) +
    geom_point() +
    facet_wrap("Month", nrow = 1) +
    scale_color_brewer(palette = "Set2") +
    labs(x = "Wind(miles/hr)", y = "Ozone (ppb)") +
    theme_bw(base_size = 15) +
    theme(legend.position = "none")

# Displaying the plot as a plotly interactive object
ggplotly(plot2)
```

## **Trend of Ozone vs Temperature in NY**

```{r plot3, echo=FALSE, cache=TRUE, fig.width=8, fig.height=5, fig.align='left', warning=FALSE, message=FALSE, dependson=c("loadData", "transformData")}
# Creating the plot with the ggplot
plot3 <- ggplot(dat, aes(Temp, Ozone, color = Month)) +
    geom_point() +
    facet_wrap("Month", nrow = 1) +
    scale_color_brewer(palette = "Set2") +
    labs(x = "Temperature(F)", y = "Ozone (ppb)") +
    theme_bw(base_size = 15) +
    theme(legend.position = "none")

# Displaying the plot as a plotly interactive object
ggplotly(plot3)
```

## **Conclusions**

- There is a greater association of ozone levels to solar radiation and both tend to increase during July-September period
- There is weak association of ozone levels to temperature of the day and both tend to increase during July-September period
- There seems to be no association between ozone levels and wind speed and wind speeds remain fairly similar across the months
  
All of these points will need to be confirmed by some kind of statistical tests, either by Student's t-test or building a linear regression model
  
## **Appendix**

### **Source**
The data were obtained from the New York State Department of Conservation (ozone data) and the National Weather Service (meteorological data)

### **References**
Chambers, J. M., Cleveland, W. S., Kleiner, B. and Tukey, P. A. (1983) *Graphical Methods for Data Analysis* Belmont, CA: Wadsworth.

### **R markdown details**
  
Written in **Rmarkdown file in `r R.version.string` using RStudio IDE**  
**Packages** used,  
  
- **datasets** : *Version `r packageVersion("datasets")`*
- **impute** : *Version `r packageVersion("impute")`*
- **ggplot2** : *Version `r packageVersion("ggplot2")`*
- **plotly** : *Version `r packageVersion("plotly")`*
  
